---
title: "Adam Fire and Jake Zukaitis Project 1.1"
output: html_notebook
---

```{r}
library(ggplot2)
library(dplyr)
```
# a
$$f(x) = \left\{\begin{array}{ll}\sin{x}, & 0 < x <a\\\sin{a}, & a<x<\pi-a\\\sin{x}, & \pi-a<x<\pi\end{array}\right.$$
```{r}
a <- 1.5
x <- seq(-pi, pi, 0.01)
f <- (x < 0 & x < a) * sin(x) +
      (x > a & x < pi-a) * sin(a) +
      (x > pi-a & x < pi) * sin(x)
plot(x, f)
```
This is a graph of f(x) on interval $-\pi<x<\pi$

# b
Note that $$b_{2k} = \frac{2}{\pi} \int_{0}^{\pi} f(x) \sin{(2kx)} \,dx = \frac{2}{\pi} (\int_{0}^{a} f(x) \sin{(2kx)} \,dx + \int_{a}^{\pi - a} f(x) \sin{(2kx)} \,dx + v)$$ 
We will consider each of these integrals seperately. The first integral is 

$$\int_{0}^{a} f(x) \sin{(2kx)} \,dx = (\frac{\sin{((2k - 1)x)}}{2(2k - 1)} - \frac{\sin{((2k + 1)x)}}{2(2k + 1)})\Bigr\rvert_{0}^{a} = \frac{\sin{((2k - 1)a)}}{2(2k - 1)} - \frac{\sin{((2k + 1)a)}}{2(2k + 1)}$$
We will leave it like this. 
For the next integral, the $\sin{a}$ term will be ignored as it is just a constant multiple. 
$$\int_{a}^{\pi - a}  \sin{(2kx)} \,dx =  \frac{\cos{(2ka)}}{2k} - \frac{\cos{(2k\pi - 2ka)}}{2k}$$
Using the identity of $\cos{(a + b)}$, $\cos{(2k\pi - 2ka)} = \cos{(2ka)}$. Substituting this into the expression for the intagral gives
$$\frac{\cos{(2ka)}}{2k} - \frac{\cos{(2k\pi - 2ka)}}{2k} = \frac{\cos{(2ka)}}{2k} - \frac{\cos{(2ka)}}{2k} = 0$$
Multiplying this by the constant $\sin{a}$ still gives $0$, so this integral is $0$.
The third integral will be 
$$\int_{\pi -a}^{\pi} \sin{(x)} \sin{(2kx)} \,dx =  (\frac{\sin{((2k - 1)x)}}{2(2k - 1)} - \frac{\sin{((2k + 1)x)}}{2(2k + 1)})\Bigr\rvert_{\pi - a}^{\pi} = \frac{\sin{((2k + 1)(\pi - a))}}{2(2k + 1)} - \frac{\sin{((2k - 1)(\pi - a))}}{2(2k - 1)}$$
Note that $(2k + 1)(\pi - a) = \pi(2k + 1) - a(2k + 1)$. Using the identity of $\sin{(a - b)}$, $\sin{(\pi(2k + 1) - a(2k + 1))} = \sin{((2k + 1)a)}$. Similarly it can be shown that $\sin{((2k - 1)(\pi - a))} = \sin{((2k - 1)a)}$. Substituting this into the expression for the intagral gives
$$\frac{\sin{((2k + 1)(\pi - a))}}{2(2k + 1)} - \frac{\sin{((2k - 1)(\pi - a))}}{2(2k - 1)} = \frac{\sin{((2k + 1)a)}}{2(2k + 1)} - \frac{\sin{((2k - 1)a}}{2(2k - 1)}$$
Summing the $3$ individual integrals togehter gives 
$$b_{2k} = \frac{2}{\pi} \int_{0}^{\pi} f(x) \sin{(2kx)} \,dx = \frac{2}{\pi}(\frac{\sin{((2k - 1)a)}}{2(2k - 1)} - \frac{\sin{((2k + 1)a)}}{2(2k + 1)} + 0 + \frac{\sin{((2k + 1)a)}}{2(2k + 1)} - \frac{\sin{((2k - 1)a}}{2(2k - 1)}) = 0$$
Consider the integral
$$ \int_{0}^{\pi} f(x) \sin{((2k+1)x)} \,dx = \int_{0}^{\pi / 2} f(x) \sin{((2k + 1)x)} \,dx + \int_{\pi / 2}^{\pi} f(x) \sin{(nx)} \,dx$$
The first integral can be written as 
$$\int_{0}^{\pi / 2} f(x) \sin{((2k + 1)x)} \,dx = \int_{0}^{a} \sin{(x)} \sin{((2k + 1)x)} \,dx + \int_{a}^{\pi / 2} \sin{(a)} \sin{((2k + 1)x)} \,dx$$
The integrals here can be expressed as follows using identities for $\sin{(a + b)}$ and $\cos{(a + b)}$
$$\int_{0}^{a} \sin{(x)} \sin{((2k + 1)x)} \,dx = (\frac{\sin{((2k - 1)x)}}{2(2k - 1)} - \frac{\sin{((2k + 1)x)}}{2(2k + 1)})\Bigr\rvert_{0}^{a} = \frac{\sin{((2k - 1)a)}}{2(2k - 1)} - \frac{\sin{((2k + 1)a)}}{2(2k + 1)}$$
$$= \frac{\sin{((2k+1)a)}\cos{(a)} - \sin{(a)}\cos{((2k+1)a)}}{2(2k-1)} - \frac{\sin{((2k+1)a)}\cos{(a)} + \sin{(a)}\cos{((2k+1)a)}}{2(2k+1)}$$
$$\int_{a}^{\pi / 2} \sin{(a)} \sin{((2k + 1)x)} \,dx = \sin{(a)}\int_{a}^{\pi / 2} \sin{((2k + 1)x)} \,dx = \sin{(a)}(\frac{-\cos{((2k+1)x)}}{2k+1})\Bigr\rvert_{a}^{\pi/2} = \frac{\sin{(a)}\cos{((2k+1)a)}}{2k+1}$$
Now looking at the integral from $\pi/2$ to $\pi$ gives
$$\int_{\pi / 2}^{\pi} f(x) \sin{(nx)} \,dx = \int_{\pi / 2}^{\pi-a}  \sin{(a)}\sin{(nx)} \,dx + \int_{\pi -a}^{\pi}  \sin{(x)}\sin{((2k+1)x)} \,dx$$
The integrals here can be expressed as follows using identities for $\sin{(a + b)}$ and $\cos{(a + b)}$
$$\int_{\pi / 2}^{\pi-a}  \sin{(a)}\sin{(nx)} \,dx = \sin{(a)}\int_{\pi/2}^{\pi -a} \sin{((2k + 1)x)} \,dx = \sin{(a)}(\frac{-\cos{((2k+1)x)}}{2k+1})\Bigr\rvert_{\pi/2}^{\pi-a} = \frac{\sin{(a)}\cos{((2k+1)a)}}{2k+1}$$
$$\int_{\pi -a}^{\pi}  \sin{(x)}\sin{((2k+1)x)} \,dx = (\frac{\sin{((2k - 1)x)}}{2(2k - 1)} - \frac{\sin{((2k + 1)x)}}{2(2k + 1)})\Bigr\rvert_{\pi-a}^{\pi} = \frac{-\sin{((2k - 1)(\pi-a))}}{2(2k - 1)} + \frac{\sin{((2k + 1)(\pi-a))}}{2(2k + 1)}$$
$$ = \frac{\sin{((2k+1)a)}\cos{(a)} - \sin{(a)}\cos{((2k+1)a)}}{2(2k-1)} - \frac{\sin{((2k+1)a)}\cos{(a)} + \sin{(a)}\cos{((2k+1)a)}}{2(2k+1)}$$
Note that both of the original integrals
$$ \int_{0}^{\pi} f(x) \sin{((2k+1)x)} \,dx = \int_{0}^{\pi / 2} f(x) \sin{((2k + 1)x)} \,dx + \int_{\pi / 2}^{\pi} f(x) \sin{(nx)} \,dx$$
are equal to each other. This means that 
$$ \int_{0}^{\pi} f(x) \sin{((2k+1)x)} \,dx = 2\int_{0}^{\pi / 2} f(x) \sin{((2k + 1)x)} \,dx$$
Plugging this into the equation for $b_n$ (using the index of $2k+1$ since this is only for odd indices) gives 
$$b_{2k+1} = \frac{4}{\pi}\int_{0}^{\pi / 2} f(x) \sin{(nx)} \,dx$$

# c
We integrate using the formula from part b using $a=\frac{\pi}{4}$.
\begin{align*}
\frac{4}{\pi}(\int_{0}^{\frac{\pi}{4}}\sin{x}\sin{(2k+1)x}\,dx+\int_{\frac{\pi}{4}}^{\frac{\pi}{2}}\sin{a}\sin{(2k+1)x}\,dx)&=\frac{4}{\pi}(\frac{\sin{2kx}}{4k}\Bigr\rvert_{0}^{\frac{\pi}{4}}-\frac{\sin{(2k+2)x}}{2k+2}\Bigr\rvert_{0}^{\frac{\pi}{4}}+\sin{a}(-\frac{2\sqrt{2}}{\pi(2k+1)}\cos{(2k+1)x}\Bigr\rvert_{\frac{\pi}{4}}^{\frac{\pi}{2}}))\\
&=\frac{1}{\pi}(\frac{\sin{\frac{\pi k}{2}}}{k}-\frac{\sin{(\frac{\pi k}{2}+\frac{\pi}{2}})}{k})+\frac{2}{\pi(2k+1)}\cos{(\frac{\pi k}{2}+\frac{\pi}{4}})
\end{align*}

# d
We integrate using the $b_n$ when $n=1$ to avoid dividing by $k=0$.
\begin{align*}
\frac{4}{\pi}(\int_{0}^{\frac{\pi}{4}}\sin^2{x}\,dx+\int_{\frac{\pi}{4}}^{\frac{\pi}{2}}\sin{a}\sin{x}\,dx)&=\frac{4}{\pi}(\frac{1}{2}(x-\frac{\cos{2x}}{2}\Bigr\rvert_{0}^{\frac{\pi}{4}})-\sin{a}\cos{x}\Bigr\rvert_{\frac{\pi}{4}}^{\frac{\pi}{2}})\\
&=\frac{1}{2}+\frac{1}{\pi}
\end{align*}

# e

```{r}
x <- seq(0, pi, length.out = 100)
x <- x[2:length(x)]
```

```{r}
X <- seq(-pi, pi, length.out = length(f))
f.df <- data.frame(x = X, f = f)
ggplot(f.df, aes(x, f))+geom_line(color="blue")
```

```{r}
fseries <- function(x,n){
  a0 = rep(0, length(x))  # a0 is direct term
  fhat <- a0
  for(i in 1:n){ # Do the following n times -- could theoretically go to infinity
    an = 0 # For this problem an is always 0
    bn = ((sin(pi/2))/i-(sin((pi/2)*i+(pi/2)))/i)/pi+(2/(pi*(2*i+1)))*cos((pi/2)*i+(pi/4)) # the code for these b_ns
    for(k in 1:length(x)){ # move through the independent variable 
      fhat[k] = fhat[k]+an*cos(i*x[k])+bn*sin(2*i*x[k]+1)
    }
  }
  return(fhat)
}
```
We do the Fourier series function using $2k+1$ as opposed to $n$ so that we can use our coefficient from part c.

```{r}
fhat <- fseries(X, 81)
f.df <- f.df %>% mutate(fhat = fhat)
ggplot(f.df)+geom_point(aes(x,fhat), color = "magenta")
```
This is the graph of the sum of the series that includes all terms until the $\sin{81x}$ term.